---
title: "Problem Set 7"
author: "Demonstration"
format: 
        html:
                self-contained: true
                code-fold: true
                code-summary: "Show the code"
editor: source
---

The preferred format would be to do your work in a code chunk at the end of this document after the instructions. Use echo=TRUE for your code chunk so that your code is visible. 

Do not alter the top lines between --- and --- except to *carefully* add your name inside the quotes.

Using your project data set, or any built in R dataset, do the following:

1. Using your chosen y (dependent) variable and one of your chosen x (explanatory) variables, perform a simple OLS regression using R. Follow these steps:

A. Load your data into a dataframe object
B. define a model object using the lm() function

        - The format is `model <- lm(y ~ x, data = dataframe)`
        - where:
        - y is the dependent variable name from your dataframe
        - x is the explanatory variable name from your dataframe
        - and dataframe is the name of your dataframe object

C. report the model statistics using the summary() function

        - The format is `summary(model)`

D. produce a model table using the stargazer() function from the stargazer library with type = "text" 

        - The format is `stargazer(model, type = "text")`
  
2. Using the same variables, produce a scatterplot with an lm() line and 95% confidence interval. You may use the base r plot() function or the ggplot() function from Lecture 15.



Load the data:


```{r}
#load demonstration_data.csv - you will use your own data and assign your own dataframe name
demonstration_data <- read.csv("demonstration_data.csv")
head(demonstration_data)

```

Define the model object using the lm function:


```{r}

#define model object using lm()
model_object1 <- lm(v2clacjstw ~ v2x_partipdem, data = demonstration_data)

```

Report the model statistics using the summary() function:

```{r}

#report model statistics using summary()

summary(model_object1)


```

Produce a model table using the stargazer() function from the stargazer library with type = "text":

```{r}

# produce a model table using stargazer()
#first it is necessary to load the stargazer package with the library command
#install.packages("stargazer") # If you have not previously installed stargazer, you will need to do this first
library(stargazer)

stargazer(model_object1, type = "text")



```
## Not from Problem Set but important

What is the equation for the regression line?

From the stargazer table:

Constant or alpha is -0.920

The coefficient for X (v2x_partipdem) is 5.167

*so the equation is*

v2clacjstw = -0.920 + 5.167* v2x_partipdem + error

Are the results significant?

The coefficient of v2x_partipdem is 5.167 and the standard error is 0.304. The t-value is 17.016. The p-value is < .05. The coefficient is significant at the 95% confidence level.

The Constant is also significant. 


Use the ggplot function to produce a scatter plot with an lm() line and 95% confidence interval for the variables from the model object:

```{r}
# first need to load ggplot2 library
#install.packages("ggplot2") # If you have not previously installed ggplot2, you will need to do this first
library(ggplot2)


ggplot(data = demonstration_data, aes(x = v2x_partipdem, y = v2clacjstw)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Confidence Interval for Regression Line",
       subtitle = "95% Confidence Interval for the Slope",
       caption = "Source: VDem ver 13") + 
        xlab("Participatory Democracy Index") + 
        ylab("Access to Justice for Women") 


```
