---
title: "Problem Set 4"
format: 
        html:
                self-contained: true
---


**For these questions, use your project data set that you have chosen for the Project assignment or any test data set other than those used in class lectures or labs.**

**This will be the first part of your semester project.**

**If you choose not to use Quarto or R Markdown, your results need to be similarly formatted to the output of a rendered Quarto or knitted R Markdown document. Readability and format matter.**

<!-- **If you choose, you may also simulate a data set with at least 100 observations of at least two variables. If you do this, give the variables descriptive column names.** -->

```{r}

sports_car_prices <- read.csv("Sport car price (1).csv")
head(sports_car_prices)


```

## Data cleanup

1. Subset to a single year so we don't have to account for age of car's effect on price
2. Rename the variables to make them a little cleaner
3, Remove the commas from price variable and coerce to numerice
4. Coerce variables to numeric
5. Drop unneeded varaibles
6 Drop unneeded variables - There are two stubborn observations in Torque. It may be necessary to use "na.rm = TRUE" with anything that involves Torque. See the example with var()
7. Save with new file name - clean_sports_car_prices.csv

```{r}

# Subset to a single year using the which(function)
sports_car_prices <- sports_car_prices[which(sports_car_prices$Year == "2021"),]



# Create new "snake case" variables for consistency
sports_car_prices$Car_Make <- sports_car_prices$Car.Make
sports_car_prices$Car_Model <- sports_car_prices$Car.Model
sports_car_prices$Torque_Ft_Lb <-sports_car_prices$Torque..lb.ft.
sports_car_prices$Price_Dollars <-sports_car_prices$Price..in.USD.

# remove the commas from Price_Dollars and change price from character to numeric data (coerce to numeric)
sports_car_prices$Price_Dollars <- as.numeric(gsub(",","",sports_car_prices$Price_Dollars))

# remove the commas from Torque_Ft_Lb and change price from character to numeric data (coerce to numeric)
sports_car_prices$Torque_Ft_Lb <- as.numeric(gsub("N/A","999999999999",sports_car_prices$Torque_Ft_Lb))

# subset to just the variables we want

sports_car_prices <- sports_car_prices[c("Car_Make","Car_Model","Price_Dollars","Torque_Ft_Lb")]

# drop NAs for Torque
sports_car_prices[which(sports_car_prices$Torque_Ft_Lb != 999999999999),]

# check the data
summary(sports_car_prices$Year)
var(sports_car_prices$Torque_Ft_Lb, na.rm = TRUE)
sd(sports_car_prices$Price_Dollars)

# write to a new CSV file
write.csv(sports_car_prices, file = "clean_sports_car_prices.csv")


```



```{r}




```

